---
title: "Accuracy of the Metropolis-Hasting Algorithm"
author: "Wenjing Xu"
date: "October 13, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This report is about the accuracy of [the Metropolis-Hasting algorithm](https://theoreticalecology.wordpress.com/2010/09/17/metropolis-hastings-mcmc-in-r/) in finding a value of *a*, for different numbers of iterations. 

### 1st Step: Background Setting
Here we assume a linear relationship between the predictor and the response variable. 

Specifically, let's assume $Y = 5X + 0 + \epsilon$, where $\epsilon$ ~ $N(0, 10^2)$

The test data used to fit our model is created as follow.
```{r}
trueA <- 5
trueB <- 0
trueSd <- 10
sampleSize <- 31

x <- (-(sampleSize-1)/2):((sampleSize-1)/2)
y <- trueA*x + trueB + rnorm(n = sampleSize, mean = 0, sd = trueSd)
```

### 2nd Step: Calculate mean and std of *a* in each resulted value of chain
The code for computing mean and std of a values in chain is as follows. 
```{r}
source("C:\\Users\\xwj93\\Documents\\GitHub\\assignment-2-wendy182\\all the functions.R")

compare_outcomes <- function(iterations){
  mean_sd_of_a = array(dim = c(10, 2))
  for (j in 1:10){
    startvalue = c()
    startvalue[1] = runif(1, min=0, max=10)
    startvalue[2] = rnorm(1, sd=5)
    startvalue[3] = runif(1, min=0, max=30)
    chain = array(dim = c(iterations+1, 3))
    chain[1, ] = startvalue
    for (i in 1:iterations){
      proposal = proposalfunction(chain[i, ])
      probab = exp(posterior(proposal) - posterior(chain[i, ]))
      if (runif(1) < probab){
        chain[i+1, ] = proposal 
      } else{
        chain[i+1, ] = chain[i, ]
      }
    }
    mean_sd_of_a[j,] = c(mean(chain[, 1]), sd(chain[, 1]))
  }
  return (mean_sd_of_a)
}
```

### 3rd Step: Compare accuray of estimation of *a* under different numbers of iterations. 
In this step, I directly call the compare_outcomes functions while giving different values of iterations. 
the reported mean and std of *a* under each scenario can help to investigate accuracy of this algorithm in estimating *a*. 
```{r}
compare_outcomes(1000)
compare_outcomes(10000)
compare_outcomes(100000)
```
The results above shows the mean and std of *a* in each of 10 loop under different values of iterations (1000,10000,100000). The first column is mean value of *a* in chain, and the second is the std of *a* in chain. 

Comapring these three matrix, we can see that as iteration times increase, the value of *a* becomes more stable which is indicated by a slower std, and its value stabilises close to the true value of a (i.e. 5). In other words, when iterate for significantly large times, this algorithm accuracy increases with the increase of iteration times. 

